---
title: "Covid91 vaccine study Final2024"
author: "Aylin Torres"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

Question: how effective is the vaccine on males?

# Method

$h_0$ the vaccine is not effected in preventing covid 91 in males 
$h_0$ the vaccine is effective in preventing covid 91 in males


Infected~Sex
Categorical ~ Categorical
We will use a helper file Cat~Cat

# Graphical 

```{r}
Male <- FinalDat[FinalDat$sex=="M",]
```

```{r}
barchartGC(~infected + treatment,data=Male)
barchartGC(~infected + treatment,data=Male, type="percent")
```

The bar chart shows that males who received the vaccine are less likely to cantact covid covid 91

#Numerical

```{r}
table1 <- xtabs(~infected + treatment, data=Male)
rowPerc(table1)
colPerc(table1)
```
The numerical data shows that within the placebo group 73.80% of males got infected with covid 911, while of the males who received the vaccine only 26.20% contracted covid91. Therefore this indicates that males who received the vaccine significantly educed the risk of contracting covid 91.


##Inferential
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)   
```

The Chi square p value is 0.000000000000022 fisher exacted tesr shows the p value is .0000000000000022 which is lower than .05 so we rejected the null.

##Conclusion

The Chi-squared statistic is 77.867, with 1 degree of freedom, and the p-value is < 2.2e-16, which is extremely small. Since the p-value is far below the common significance level of 0.05, we reject the null hypothesis. The null hypothesis stated that there is no association between the treatment (placebo vs. vaccine) and the infection status (COVID-19 vs. OK).the vaccine group has a significantly lower infection rate (1.02%) compared to the placebo group (2.82%). This strongly suggests that the vaccine reduces the likelihood of COVID-19 infection compared to the placebo treatment.

## Females

Question: How effective is the vsccine on females?

##Methodsn

$h_0$ the vaccine is not efected in preventing covid 91in females 
$h_A$ the vaccine is effective on preventing covid91 in the females 

Infected~Sex
Categorical ~ Categorical
We will use a helper file CAt~Cat

##Graphical

```{r}
Female <- FinalDat[FinalDat$sex=="F",]  
```

```{r}
barchartGC(~infected + treatment,data =Female)
barchartGC(~infected + treatment,data =Female, type= "percent")

```
The bar chart shows that female that received the vaccine are less likely to contract covid 91

##Numerical

```{r}
tabel1 <- xtabs(~infected + treatment, data=Female)
rowPerc(table1)
colPerc(table1)
```

The numerical data shows that within the placboo group 64.93% of females got infected with covid 92, while of the females who received the vaccine only 35.07% contracted covid92. Therefore this indicates that the females who received the vaccine significantly reduced the risk of contracting covid91.

#Inferential

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

the chi square test shows the p value .00000000000022 and the fisher exact test shows the p value is .000000000022 which is lower than 0.5 so we reject the null hypothesis 


##Conclusion

The bar chart shows that females that received the vaccine are less likely to contract covid 91. The numerical data shows that within the placebo group 2.82% of the females got infected with covid91. Therefore this indicates that the females who received the vaccine significantly reduced the risk of contracting covid91. The fisher exact test shows the p-value is .0000000000022 which is lower than 0.5 so we reject the null hypothesis.


## LGBTQ
Question: How effective is the vaccine on LGBTQ?

##Method 

$H_0$ the vaccine is not effective in preventing covid91 in LGBTQ
$H_0$ the vaccine is effective in preventing covid 91 in LGBTQ

Infected~LGBTQ(gay)
Categorical ~ Categorical
We will use a helper file Cat~Cat

##Graphical

```{r}
LBGTQ <- FinalDat[FinalDat$LGBTQ=='gay',]
```

```{r}
barchartGC(~infected + treatment,data=LBGTQ)
barchartGC(~infected + treatment,data=LBGTQ, type="percent" )
```

the bar chart shows that members if the LBGTQ community are less likely to contract covid 91 if your recived the vaccine.

##Numerical

```{r}
table1 <- xtabs(~infected + treatment, data=LBGTQ)
rowPerc(table1)
colPerc(table1)  
```
The numerical data shows that within the placebo group 80% of the LBGTQ
community members got infected with covid91, while of the LBGTQ community members who received the vaccine only 20% contracted covid91.

##Inferential

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

the chi square test p value is.0000000002266 and the fisher exact test shows the p value is .0000007269 which is lower than 0.05 so we reject the null.

##Conclusion

The bar chart shows that members if the LBGTQ community are less likely to contract covid 91 if your received the vaccine. The numerical data shows that within the placebo group 805 of tthe LBGTQ community members got infected with covid 91 while of the LBGTQ community members who receid the vaccine only 20% contracted covid 91. The fisher exact test shows the p value is .0000000007269 which is lower than .05 so we rejected the null hypothesis.

## Druggies

Question: How effective is the vaccine on druggies?

##Methods 

$H_0$  The vaccine is not effective in preventing covid 91 in druggies
$H_0$ The vaccine is effective in preventing covid 91 in druggies

Infected~DrugUser
Categorical ~ Categorical
We will use a helper file Cat~Cat



##graphical Results


```{r}
DrugUser <- FinalDat[FinalDat$DrugUser=='yes',]
```

```{r}
barchartGC(~infected + treatment,data=DrugUser)
barchartGC(~infected + treatment,data=DrugUser, type="percent")

```

According to the barchart you are more likely to get covid91 if receive the vaccine as a drug user.


##Numerical Results


```{r}
table1 <- xtabs(~infected + treatment, data=DrugUser)
rowPerc(table1)
colPerc(table1)
```

The numerical data shows the results in percentages.Within the placebo group 8.33% of ndividuals got infected with covid91. While in the vaccine group 91.67% contracted covid91. This indicates that receiving the vaccine as a drug usesr significantly increase the risk of contracted covid91.


##inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The chi-square test shows a p-value of 1.414e-10 data and the fisher exact test shoes that the p-value is 3.261e-12 which is significantly lower than 0.05 which means we reject the null hypothesis.


##conclusion
The bar chart shows that you are far more likely to contract covid 91 if you did receive the vaccine as a drug user.The numerical data shows the results in percentages. within the placebo group 8.33% of individuals got infected with covid91. While in the vaccine group 91.67% contracted covid 91.The inferential data shows a chia-squared p-value of 1.414e-10 and the data includes a fisher exact test showing that the p-value is 3.261e-12 which is significantly lower than 0.05 which means we reject the null hypothesis.In summation you receive the vaccine



## MAle conclusion
When trying to figure out the effectiveness of the vaccine in males,the bar chart shows that you are less likely to contract covid91 if you did receive the vaccine. The numerical data shows the results in percentages. Within the placebo group 73.80% of individuals got infected with covid91. While in the vaccine group 26.20% contracted covid-91. This indicates that receiving the vaccine as a male significantly reduces the risk of contracted covid-91.Also shows the Fisher exact test that shows that the P-Value is 2.2-16 which is significantly shows that is as aa male you are far less likely to contract covid91 upon receing the vaccine.

##Female Conclusion

When testing the effectiveness of the vaccine in females,
the bar chart shows that you are less likely to
contract Covid91 if you did receive the vaccine. The numerical data shows the results in percentages. Within the placebo group 64.93% of individuals got infected with Covid91.While in the vaccine group 35.07% contracted covid-91. This indicates that receiving the vaccine as a female significantly reduces the risk of contracted Covid-91. The inferential data shows the Chi-Square P-Value of 1.914e-08, and the data also includes the Fisher exact test that shows that the P-Value is 1.246-08 which is significantly lower than 0.05 which means we reject the null hypothesis. In summation these examples and simulations show that is as a female you are far less likely to contract covid-91 if you receive the vaccine.

##LGBTQ Conclusion

When testing the effectiveness of the vaccine on members of the LGBTQ community, the bar chart shows that you are less likely to contract Covid91 if you did receive the vaccine. The numerical data shows the results in percentages. Within the placebo group 80.00% of individuals got infected with Covid91. While in the vaccine group 26.00 contracted covid-91. This indicates that receiving the vaccine as a LGBTQ member significantly reduces the risk of contracted Covid-91. The inferential data shows that the Chi-Squared P-Value is 2 266e-06, and the data also includes the Fisher exact test that shows that the P-Value is 7.269e-07 which is significantly lower than 0.05 which means we reject the null hypothesis. In summation these examples and simulations show that is as a member of the LGBTQ community you are far less likely to contract covid-91 if you receive the vaccine.


##Drug user conclusion 

When testing the effectiveness of the vaccine on people who are drug users, the bar chart shows that you are far more likely to contract Covid91 if you did receive the vaccine as a drug user.
The numerical data shows
the results in percentages. Within the placebo group 8.33% of individuals got infected with Covid91. While in the vaccine group 91.67% contracted covid-91. This indicates that receiving the vaccine as a drug user significantly increases the risk of contracted Covid-91. The inferential data shows that the Chi-Squared and the data also includes the Fisher exact test that shows that the p-Value is 1.414e-10 and the data also includes the fishers exact test that shows that p-value is 3.261e-12 which is significantly Lower than 0.05 which means we reject the null hypothesis. In summation these examples and simulations show that is as a drug user you are far more likely to contract covid-91 if you receive the vaccine.


##Overall Conclusion 

Overall, upon reviewing the data it is clear that receiving the vaccine will always have an effect on the person receiving it.
The data clear shows that receiving the vaccine is generally effective in preventing the
contraction of Covid-91 if you are male, female, or a member of the LGBT community.However, the opposite is true for people who are drug users.According the studies conducted you are at a much higher risk of contracting Covid-91 if you are a drug user. Presumably the opposite effect of the vaccine's purpose.




















# Overall Results and Conclusions